---
title: "CBE1"
format: pdf
---

# Introduction:
- Comparison of presidential debates and how they have changed over the past decade
- Changes over time for parties, incumbents, moderators, etc.
- Types of words used, amount of words spoken, other differences

# Data:
- First presidential debates for Obama vs Romney and Harris vs Trump
- Obtained transcripts online and split them by who was speaking
- 3 groups for each transcript: Democrat, Republican, and Moderator

```{r, include=FALSE}
library(quanteda)
library(quanteda.textstats)
library(tidyverse)
library(gt)
source("../R/helper_functions.R")
```

```{r, include=FALSE}
obama_txt <- readLines("obama.txt")
obama_freq <- obama_txt %>%
corpus() %>%
tokens(what = "word", remove_punct = TRUE) %>%
dfm() %>%
textstat_frequency() %>%
as_tibble() %>%
dplyr::select(feature, frequency) %>%
rename(Token = feature, AF = frequency) %>%
mutate(RF = AF/sum(AF)*100)
```

```{r, echo=FALSE}
head(obama_freq)
```

# Methods:
You should clearly explain which methods you chose and why.


# Results:
The results should begin with descriptive analysis or EDA (exploratory data analysis) before moving on to other kinds of analysis/modeling. Be sure your visualizations are well-designed and properly captioned. Also, be sure to follow the appropriate conventions for reporting your statistical results as outlined in Brezina.

# Discussion:
What conclusions might you draw from your findings? Do they suggest interesting directions for further study? If your CBE didn’t work out the way you had hoped/expected, do you have any intuitions as to why? What are the limitations of your study and approach?
